Computable Poincar\'e--Friedrichs constants\\ for the $L^{p}$ de Rham complex over domains\\ with shellable triangulations

### 2024-07-15 main.tex
remaining annotations

We return to the theory of triangulations,
as our main objective requires some further concepts.
We are interested in simplicial complexes that triangulated manifolds. 
We put particular focus on \emph{shellable} simplicial complexes as an important class of complexes. Such simplicial complexes are constructed by successively adding simplices in a particularly well-behaved manner. Patches within triangulations of dimension two and three are examples of such shellable complexes. 
Our main references for this section are the monographs by Kozlov~\cite{kozlov2008combinatorial} and Ziegler~\cite{ziegler2012lectures}.

@@

We define an $n$-dimensional simplicial complex to be a \emph{manifold triangulation} if the underlying set $\underlying{\calT}$ is an $n$-dimensional manifold with boundary.
We recall that this means that for every $x \in \underlying{\calT}$
there exists $\epsilon > 0$ and an embedding $\phi : \Ball_\epsilon(x) \cap \underlying{\calT} \rightarrow \bbR^{n}$
such that $\phi(0) = 0$ and $\phi$ is an isomorphism either onto the closed unit ball $\calB = \{ x \in \bbR^{n} \suchthat |x| \leq 1 \}$
or onto $\{ x \in \calB \suchthat x_1 \geq 0 \}$.
Any simplicial complex that triangulates an $n$-dimensional manifold must be $n$-dimensional.

@@

\begin{lemma}\label{lemma:boundarysimplices}
    Let $\calT$ be a finite $n$-dimensional simplicial complex whose underlying space is a manifold $\calM$. 
    \begin{itemize}
        \item Any face $F \in \Faces(\calT)$ is not contained in the boundary if and only if it is contained in exactly two $n$-simplices.
        \item Any face $F \in \Faces(\calT)$ is contained in the boundary if and only if it is contained in exactly one $n$-simplex.
        \item The simplices contained in the boundary constitute a triangulation of the boundary.
        % \item $\calT$ is non-branching, that is, every $F \in \calT$ of dimension $n-1$ is contained in at most two $n$-simplices of $\calT$. 
    \end{itemize}
\end{lemma}
\begin{proof}
    We prove these statements in several steps.
    \begin{enumerate}
    \item 
    We $\mathring \calM := \calM \setminus \partial \calM$ denote the interior of the manifold. 
    We will use the following fact: if $S \in \calT$ has an inner point that lies on $\partial \calM$, then all inner points of $S$ are on $\partial \calM$. 
    Since the boundary is closed, every $S \in \calT$ is either a subset of the boundary or all its inner points lie in the interior $\mathring \calM$ of the manifold.

    \item
    Let $F \in \Faces(T)$ and let $z_F \in F$ be its barycenter. 
    Let $\mathring B_F$ be an open neighborhood around $z_F$ so small that it only intersects those $n$-simplices of $\calT$ that already contain $z_F$ and no faces other than $F$.
    Suppose there are distinct $n$-simplices $T_1, T_2, \dots, T_K$ that contain $z_F$. 
    The intersection of any two of them is $F$, and their interiors are disjoint.

    If $z_F$ is an interior point of $\calM$, 
    then we can assume that $\mathring B_F$ is homeomorphic to an open $n$-ball and $\partial \mathring B_F$ is homeomorphic to a sphere $S$ of dimension $n-1$. 
    At the same time, $X = F \cap \partial \mathring B_F$ is empty if $n=1$ or homeomorphic to a sphere $S' \subseteq S$ of dimension $n-2$. % TODO: proofread
    We see that $\partial \mathring B_F \setminus X$ has $K$ distinct connected components. 
    But by construction $\partial \mathring B_F \setminus X$ is homeomorphic to $S \setminus S'$,
    which has exactly two distinct connected components according to the Jordan-Brouwer separation theorem.
    We conclude that $F$ is contained in two $n$-simplices of $\calT$.
    If an inner face $F$ were contained in only one $n$-simplex,
    then we could construct a homeomorphism from the $n$-ball onto itself that maps an interior point on the its boundary, which cannot be.
    So each inner face is contained in exactly two $n$-simplices.

    If $F$ is contained in only one $n$-simplex, then it must be a face in the boundary of $\calM$. 
    If $F$ were contained in at least two $n$-simplices,
    then we could construct a homeomorphism from the $n$-ball onto itself that maps an interior point on the its boundary, which cannot be.
    So each face within $\partial \calM$ is contained in exactly one $n$-simplex.

    \item
    Clearly, the simplices of $\calT$ contained in the boundary constitute a simplicial complex. 
    If $x \in \partial \calM$ is a boundary simplex, then it is an inner point of some simplex $S \in \calT$.
    But then this simplex must be a boundary simplex, so the boundary simplices form a triangulation of $\partial \calM$.
    \end{enumerate}
    All relevant results are proven. 
\end{proof}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% TODO: integrate into narrative 
Suppose that $\calT$ is an $n$-dimensional simplicial complex that triangulates a manifold. 
Those simplices of the manifold triangulation that are subsets of the boundary of the underlying manifold are called \emph{boundary simplices}. 
% The simplices that lie in the boundary of the underlying set are called \emph{boundary simplices}.
All other simplices of the manifold triangulation are called \emph{inner simplices}.

We have seen that the boundary simplices of a manifold triangulation constitute a triangulation of the manifolds boundary. 
We call this simplicial complex the \emph{boundary complex}. It has dimension $n-1$. 
% We see that the family of boundary faces constitute a simplicial complex, which we call \textit{boundary complex}.
% Any point on the boundary lies within some boundary simplex, and hence the boundary complex is a triangulation of the boundary. 
% Definitions imply that the boundary complex is a manifold triangulation of dimension $n-1$.
% The boundary simplices of a manifold triangulation of dimension $n$ constitute a manifold triangulation themselves, of dimension $n-1$, which triangulate the boundary of the underlying space. 
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\\

@@

\begin{remark}
It is often very helpful if a simplicial complex can be created recursively, starting with a single simplex. In particular, this enables the inductive method of proof.
Formally, a triangulation is shellable if its full-dimensional simplices can be enumerated such that each simplex intersects the union of the previously listed simplices in a codimension one triangulation. 
Roughly speaking, a triangulation is shellable if the triangulation can be built, starting from a single simplex,
by successively gluing simplices to intermediate triangulations such that a combinatorial-geometric condition is satisfied:
the gluing interface is a triangulated submanifold of the boundary. 
This forces the intermediate triangulations to be particularly well-shaped. 
\end{remark}

We turn our attention to a different idea of how to incrementally construct a simplicial complex while maintaining well-behaved intermediate complexes. 
We build upon the notion of shelling, as introduced in~\cite[Definition 8.1]{ziegler2012lectures}.
Our definition of shelling is equivalent to the notion of the shellings of simplicial complexes in the aforementioned reference; see also~\cite[Remark~8.3]{ziegler2012lectures}.

@@

\begin{lemma}    
    Suppose that an $n$-dimensional simplicial complex $\calT$ has a shelling $T_{0}, T_{1}, T_{2}, \dots, T_{M}$ and is non-branching, 
    that is, each simplex of dimension $n-1$ is contained in at most two simplices. 
    % TODO: what is non-branching?
    Then
    $$
        X_{m} := T_{0} \cup T_{1} \cup \dots \cup T_{m}
    $$ 
    is a manifold with boundary for all $0 \leq m \leq M$.
    % 
    In particular, $X_{m}$ is a topological $n$-ball when $m < M$, 
    and 
    $X_{M}$ is either a topological $n$-ball or topological $n$-sphere. 
\end{lemma}
\begin{proof}  
    We prove this claim by induction. 
    Certainly, if $\calT$ contains only one single $n$-simplex, then it is a shellable triangulation of a topological $n$-ball. 
    Suppose that $X_m := T_{0} \cup T_{1} \cup \dots \cup T_{m}$ is a topological $n$-ball and that $T_{m+1}$ is the next $n$-simplex in the shelling.
    By definition, $I_{m+1} := X_{m} \cap T_{m+1}$ is a submanifold of $\partial T_{m+1}$,
    and it is triangulated by some faces of $T_{m+1}$ and their subsimplices.

@@

    \todo{Double check this proof and make it more explicit.}
    Write $M = |\calT|$. 
    We call a triangle $T \in \calT$ \emph{good} if it intersects the boundary $\partial M$ in a non-empty union of edges. % TODO: I understand what it means from the rest of the proof, but I believe it could a bit more explicit.
    We show by an induction argument over the number of triangles that every triangulation of a $2$-disk contains a good triangle. 
    Clearly, this is the case if the triangulation contains only one triangle. 
    Now suppose the induction claim is true when the triangulation includes $N$ triangles,
    and assume that $\calT$ includes $N+1$ triangles. 
    There exists a triangle $T' \in \calT$ with at least one edge in the boundary. 
    If $T'$ is good, then we are already done. 
    Otherwise, $T'$ intersects the boundary at one edge and its opposing vertex.
    Removing $T'$ splits the manifold into two pieces, each of which is a topological $2$-disk.
    By the induction assumption, they contain a good triangle.
    \todo{MV:But what if this is the neighbor of $T'$ and adding $T'$ back makes the triangle bad?}

@@

We discuss the Sobolev spaces of vector calculus and their transformation behavior. 
Let $\Omega \subseteq \bbR^{3}$ be a bounded open set. 
First, recall that $L^{p}(\Omega)$ is the space of scalar-valued $p$-integrable functions defined on $\Omega$
and that $\bfL^{p}(\Omega) := L^{p}(\Omega)^{n}$ for vector-valued functions with each component in $L^{p}(\Omega)$. 
In the three-dimensional setting, we are particularly interested in the Sobolev vector analysis. 
$W^p(\grad,\Omega) := W^{1,p}(\Omega)$ is the space of scalar-valued $L^{p}(\Omega)$ functions with weak gradients in $\bfL^{p}(\Omega)$, that is,  
\[
    W^p(\grad,\Omega) = \{ u \in L^{p}(\Omega) \suchthat \grad v \in \bfL^{p}(\Omega) \}
\]
The space $\bfW^{p}(\curl,\Omega)$ of vector-valued $\bfL^{p}(\Omega)$ functions with weak curls in $\bfL^{p}(\Omega)$
and the space of vector-valued $\bfL^{p}(\Omega)$ functions with weak divergences in $L^{p}(\Omega)$ are defined as 
\[
    \bfW^{p}(\curl,\Omega) = \{ \bfu \in \bfL^{p}(\Omega) \suchthat \curl \bfu \in \bfL^{p}(\Omega) \},
    \quad 
    \bfW^{p}(\divergence,\Omega) = \{ \bfu \in \bfL^{p}(\Omega) \suchthat \divergence \bfu \in L^{p}(\Omega) \}.
\]
We are interested in transformations of these Sobolev tensor fields from one domain onto the other. 
Suppose that $\Omega, \Omega' \subset \mathbb{R}^3$ are open sets and suppose that $\phi: \Omega \to \Omega'$ is a bi-Lipschitz mapping.
The gradient-, curl-, divergence-conforming Piola transformations are the mappings 
$\phi^{\grad}: L^{p}(\Omega') \to L^{p}(\Omega)$,
$\phi^{\curl}: \bfL^{p}(\Omega') \to \bfL^{p}(\Omega)$, 
$\phi^{\divergence}: \bfL^{p}(\Omega') \to \bfL^{p}(\Omega)$,
and
$\phi^{\rm b}: L^{p}(\Omega') \to L^{p}(\Omega)$. 
We define them 
for any $v \in L^{p}(\Omega')$ and $\bfw \in \bfL^{p}(\Omega')$ by setting 
\begin{subequations}\label{eq_definition_piola}
\begin{align}
    \phi^{\grad}(v) &= v \circ \phi, \\
    \phi^{\curl}(\bfw) &= (\Jacobian\phi^{T} \bfw) \circ \phi, \\
    \phi^{\divergence}(\bfw) &= \left( \adj(\Jacobian\phi) \bfw \right)  \circ \phi, \\
    \phi^{\rm b}(v) &= \left( \det(\Jacobian\phi) v \right)  \circ \phi,
\end{align}
\end{subequations}
and where $\Jacobian\phi$ is the Jacobian matrix of $\phi$. 
These mappings are invertible. 
%
% TODO: bounds on the Jacobian transform 
%
We use the commutativity relations 
\begin{subequations}\label{eq_piola_commute}
\begin{align}
    \grad (\phi^{\grad} (v)) &= \phi^{\curl}(\grad v), 
    \\
    \curl (\phi^{\curl}(\bfv)) &= \phi^{\divergence}(\curl \bfv) 
    % TODO: watch the sign 
    \\
    \divergence (\phi^{\divergence}(\bfw)) &= \phi^{\rm b}(\divergence \bfw)
\end{align}
\end{subequations}
for $v \in W^{p}(\grad,\Omega)$, $\bfv \in \bfW^{p}(\curl,\Omega)$, and $\bfw \in \bfW^{p}(\divergence,\Omega)$.  
We summarize this as a commuting diagram:
\begin{align*}
    \begin{CD}
        W^{p}(\grad,\Omega') @>{\grad}>> \bfW^{p}(\curl,\Omega') @>{\curl}>> \bfW^{p}(\divergence,\Omega') @>{\divergence}>> L^{p}(\Omega')
        \\
        @VV{\phi^{\grad}}V 
        @VV{\phi^{\curl}}V 
        @VV{\phi^{\divergence}}V 
        @VV{\phi^{\rm b}}V 
        \\
        W^{p}(\grad,\Omega ) @>{\grad}>> \bfW^{p}(\curl,\Omega ) @>{\curl}>> \bfW^{p}(\divergence,\Omega ) @>{\divergence}>> L^{p}(\Omega )
    \end{CD}
\end{align*}

@@

We write $C^{\infty}\Alt^{k}(\Omega)$ for the space of smooth differential $k$-forms over $\Omega \subseteq \bbR^{n}$,
which is the vector space of smooth mappings from $\Omega$ into $\Alt^{k}(\bbR^{n})$.
The exterior derivative \( \cartan \) is an operator that takes a \( k \)-form \( \omega \in C^{\infty}\Alt^{k}(\Omega) \) 
to a \((k+1)\)-form \( \cartan\omega \in C^{\infty}\Alt^{k+1}(\Omega) \). 
Every \( k \)-form \( \omega \in C^{\infty}\Alt^{k}(\Omega) \) can be written 
\[
    \omega = 
    \sum_{ \sigma \in \Sigma(k,n) } 
    \omega_{\sigma} \, 
    \cartanx^{\sigma(1)} \wedge \cartanx^{\sigma(2)} \wedge \cdots \wedge \cartanx^{\sigma(k)},
\]
where \( \omega_{\sigma} : \Omega \rightarrow \bbR \) are smooth functions.
The exterior derivative \( \cartan\omega \) is defined by
\[
    \cartan\omega = 
    \sum_{ \sigma \in \Sigma(k,n) } 
    \sum_{j=1}^{n} 
    \frac{\partial \omega_{\sigma}}{\partial x^j} 
    \, \cartanx^j \wedge 
    \cartanx^{\sigma(1)} \wedge \cartanx^{\sigma(2)} \wedge \cdots \wedge \cartanx^{\sigma(k)}.
\]
The exterior derivative is linear and nilpotent, which means 
\( \cartan(\cartan\omega) = 0 \) for any \( \omega \in C^{\infty}\Alt^{k}(\Omega) \).
Moreover, it satisfies the Leibniz rule:
\[ 
    \cartan(\omega \wedge \eta) 
    = 
    \cartan\omega \wedge \eta + (-1)^{k} \omega \wedge \cartan\eta, 
    \qquad \omega \in C^{\infty}\Alt^{k}(\Omega), 
    \qquad \eta \in C^{\infty}\Alt^{l}(\Omega).
\]
Let us now turn our attention to Sobolev spaces of differential forms. 
Since the exterior product space $\Alt^{k}(\bbR^{n})$ carries a canonical norm, we can consider the pointwise norms of differential $k$-forms. 
% First, we notice that the % TODO: introduce a norm 
We let $L^{p}\Alt^{k}(\Omega)$ be the space of differential $k$-forms over $\Omega$ with locally integrable coefficients 
such that its pointwise norm is $p$-integrable. 
The exterior derivative is defined in the sense of distributions and we introduce 
\[
    W^p\Alt^{k}(\Omega) = \{ u \in L^{p}\Alt^{k}(\Omega) \suchthat \cartan v \in L^{p}\Alt^{k+1}(\Omega) \}
\]
We observe that $u \in L^p\Alt^{k}(\Omega)$ has weak exterior derivative $f \in L^p\Alt^{k+1}(\Omega)$
if and only if for all $v \in C^{\infty}_{c}\Lambda^{n-k-1}(\Omega)$ we have the integration by parts formula
\begin{align*}
    \int_{\Omega} \cartan v \wedge u
    &=
    (-1)^{k(n-k)+1}
    \int_{\Omega} v \wedge f 
    .
\end{align*}
We let $W^p_{0}\Alt^{k}(\Omega)$ be the closed subspace of those $u \in W^p\Alt^{k}(\Omega)$ 
for which the extension by zero $\tilde u : \bbR^{n} \rightarrow \Alt^{k}(\bbR^{n})$ is a member of $\tilde u \in W^p\Alt^{k}(\bbR^{n})$. 
The space $W^p_{0}\Alt^{k}(\Omega)$ formalizes the idea of differential forms with vanishing boundary trace along the boundary $\partial\Omega$.

@@

>>>>

We return to the theory of triangulations,
as our main objective requires some further concepts.
We are interested in simplicial complexes that triangulate manifolds. 
Our particular interest are manifold triangulations that are \emph{shellable}.
Such simplicial complexes are constructed by successively adding simplices in a particularly well-behaved manner. 
Patches within triangulations of dimension two and three are examples of such shellable complexes. 
The monographs by Kozlov~\cite{kozlov2008combinatorial} and Ziegler~\cite{ziegler2012lectures} are our main references for this section.

@@

We define an $n$-dimensional simplicial complex to be a \emph{manifold triangulation} if the underlying set $\underlying{\calT}$ is an $n$-dimensional manifold with boundary.
We recall that this means that for every $x \in \underlying{\calT}$
there exists $\epsilon > 0$ and an embedding $\phi : \Ball_\epsilon(x) \cap \underlying{\calT} \rightarrow \bbR^{n}$
such that $\phi(0) = 0$ and $\phi$ is an isomorphism either onto the closed unit ball $\calB = \{ x \in \bbR^{n} \suchthat |x| \leq 1 \}$
or onto $\{ x \in \calB \suchthat x_1 \geq 0 \}$.
In the former case, $x$ is called an \emph{interior point}, and in the latter case $x$ is called a \emph{boundary point}. 
Any simplicial complex that triangulates an $n$-dimensional manifold must be $n$-dimensional.

@@

\begin{lemma}\label{lemma:boundarysimplices}
    Let $\calT$ be a finite $n$-dimensional simplicial complex whose underlying space is a manifold $\calM$. 
    \begin{itemize}
        \item Any face $F \in \Faces(\calT)$ is not contained in the boundary if and only if it is contained in exactly two $n$-simplices.
        \item Any face $F \in \Faces(\calT)$ is contained in the boundary if and only if it is contained in exactly one $n$-simplex.
        \item The simplices contained in the boundary constitute a triangulation of the boundary.
        % \item $\calT$ is non-branching, that is, every $F \in \calT$ of dimension $n-1$ is contained in at most two $n$-simplices of $\calT$. 
    \end{itemize}
\end{lemma}
\begin{proof}
    We prove these statements in several steps.
    \begin{enumerate}
    \item 
    We $\mathring \calM := \calM \setminus \partial \calM$ denote the interior of the manifold. 
    We will use the following fact: if $S \in \calT$ has an inner point that lies on $\partial \calM$, then all inner points of $S$ are on $\partial \calM$. 
    Since the boundary $\partial \calM$ is closed, every $S \in \calT$ is either a subset of the boundary or all its inner points lie in the interior $\mathring \calM$ of the manifold.

    \item
    Let $F \in \Faces(T)$ and let $z_F \in F$ be its barycenter. 
    Since $\calT$ is finite, we let $\mathring B_F$ be an open neighborhood around $z_F$ so small that it only intersects those $n$-simplices of $\calT$ that already contain $z_F$ and no faces other than $F$.
    Suppose there are distinct $n$-simplices $T_1, T_2, \dots, T_K$ that contain $z_F$. 
    The intersection of any two of them is $F$, and their interiors are disjoint.

    If $z_F$ is an interior point of $\calM$, 
    then we can assume that $\mathring B_F$ is homeomorphic to an open $n$-ball and $\partial \mathring B_F$ is homeomorphic to a sphere $S$ of dimension $n-1$. 
    At the same time, $X = F \cap \partial \mathring B_F$ is empty if $n=1$ or homeomorphic to a sphere $S' \subseteq S$ of dimension $n-2$. % TODO: proofread
    We see that $\partial \mathring B_F \setminus X$ has $K$ distinct connected components. 
    But by construction $\partial \mathring B_F \setminus X$ is homeomorphic to $S \setminus S'$,
    which has exactly two distinct connected components according to the Jordan-Brouwer separation theorem.
    We conclude that $F$ is contained in two $n$-simplices of $\calT$.
    If an inner face $F$ were contained in only one $n$-simplex,
    then we could construct a homeomorphism from the $n$-ball onto itself that maps an interior point on the its boundary, which cannot be.
    So each inner face is contained in exactly two $n$-simplices. 
    \todo{More details}

    If $F$ is contained in only one $n$-simplex, then it must be a face in the boundary of $\calM$. 
    If $F$ were contained in at least two $n$-simplices,
    then we could construct a homeomorphism from the $n$-ball onto itself that maps an interior point on the its boundary, which cannot be.
    So each face within $\partial \calM$ is contained in exactly one $n$-simplex. 
    \todo{More details}

    \item
    Clearly, the simplices of $\calT$ contained in the boundary constitute a simplicial complex. 
    Every $x \in \calM$ is an inner point of some simplex $S \in \calT$. 
    If $x \in \partial \calM$ is a boundary point, then $S$ must be a boundary simplex, 
    so the boundary simplices triangulate all of $\partial \calM$.
    \end{enumerate}
    All desired results are thus proven. 
\end{proof}

Suppose that $\calT$ is an $n$-dimensional simplicial complex that triangulates a manifold. 
Those simplices of the manifold triangulation that are subsets of the boundary of the underlying manifold are called \emph{boundary simplices}. 
% The simplices that lie in the boundary of the underlying set are called \emph{boundary simplices}.
All other simplices of the manifold triangulation are called \emph{inner simplices}. 
We have seen that the boundary simplices of a manifold triangulation constitute a triangulation of the manifolds boundary. 
We call this simplicial complex the \emph{boundary complex}. It has dimension $n-1$. 
% We see that the family of boundary faces constitute a simplicial complex, which we call \textit{boundary complex}.
% Any point on the boundary lies within some boundary simplex, and hence the boundary complex is a triangulation of the boundary. 
% Definitions imply that the boundary complex is a manifold triangulation of dimension $n-1$.
% The boundary simplices of a manifold triangulation of dimension $n$ constitute a manifold triangulation themselves, of dimension $n-1$, which triangulate the boundary of the underlying space. 
\\

@@

The notions of \emph{shelling} and \emph{shellable triangulation} have been discussed widely in combinatorial topology and polytopal theory. 
Formally, a triangulation is shellable if its full-dimensional simplices can be enumerated such that each simplex intersects the union of the previously listed simplices in a codimension one triangulation of a manifold. 
We thus build upon the notion of shelling, as introduced in~\cite[Definition 8.1]{ziegler2012lectures},
and our definition of shelling is equivalent to the notion of the shellings of simplicial complexes in the aforementioned reference; see also~\cite[Remark~8.3]{ziegler2012lectures}.

Roughly speaking, a triangulation is shellable if the triangulation can be built, starting from a single simplex,
by successively gluing simplices to intermediate triangulations such that a combinatorial-geometric condition is satisfied:
the gluing interface is a triangulated submanifold of the boundary. 
This forces the intermediate triangulations to be particularly well-shaped. 
We turn our attention to a different idea of how to incrementally construct a simplicial complex while maintaining well-behaved intermediate complexes. 
It is often very helpful if a simplicial complex can be created recursively, starting with a single simplex, 
as this enables the inductive method of proof.
\\

@@

\begin{lemma}    
    Suppose that an $n$-dimensional simplicial complex $\calT$ has a shelling $T_{0}, T_{1}, T_{2}, \dots, T_{M}$ and is non-branching, 
    that is, each simplex of dimension $n-1$ is contained in at most two simplices. 
    % DONE: what is non-branching?
    Then
    $$
        X_{m} := T_{0} \cup T_{1} \cup \dots \cup T_{m}
    $$ 
    is a manifold with boundary for all $0 \leq m \leq M$.
    % 
    In particular, $X_{m}$ is a topological $n$-ball when $m < M$, 
    and 
    $X_{M}$ is either a topological $n$-ball or topological $n$-sphere. 
\end{lemma}
\begin{proof}  
    We prove this claim by induction. 
    Certainly, if $\calT$ contains only one single $n$-simplex, then it is a shellable triangulation of a topological $n$-ball. 
    Suppose that $X_m := T_{0} \cup T_{1} \cup \dots \cup T_{m}$ is a topological $n$-ball and that $T_{m+1}$ is the next $n$-simplex in the shelling.
    By definition, $I_{m+1} := X_{m} \cap T_{m+1}$ is a submanifold of $\partial T_{m+1}$,
    and it is triangulated by some faces of $T_{m+1}$ and their subsimplices.

@@

    %\todo{Double check this proof and make it more explicit.}
    Write $M = |\calT|$. 
    There is nothing to show if $\calT$ contains only one triangle. 
    We call a triangle $T \in \calT$ \emph{good in $\calT$} if it intersects the boundary $\partial M$ in a non-empty union of edges. 
    Hence a triangle is good in $\calT$ if its intersection with $\partial M$ is either one, two, or three edges,
    and a triangle is not good in $\calT$ if that intersection is either empty only some of its vertices, or a vertex and the opposite edge.
    % TODO: I understand what it means from the rest of the proof, but I believe it could a bit more explicit.
    We show by an induction argument over the number of triangles that every triangulation of a $2$-disk that contains at least two triangles also contains at least two good triangles.

    Clearly, this is the case if the triangulation of the $2$-disk contains two triangles. 
    Now suppose the induction claim is true when the triangulation includes at most $N$ triangles,
    and assume that $\calT$ includes $N+1$ triangles. 
    There are at least two triangles with an edge on the boundary. 
    Suppose that $\calT$ does not have at least two triangles that are good in $\calT$,
    then there exists a triangle $T'$ that intersects $\partial M$ 
    in one edge and its opposite vertex.
    Removing $T'$ splits the manifold into two strongly connected components, each of which is a topological $2$-disk.
    By the induction assumption, each of those components contains at least two triangles 
    that are good in the respective component. 
    So each component has at least one triangle that is good in $\calT$.

    % There exists a triangle $T' \in \calT$ with at least one edge in the boundary. 
    % If $T'$ is good, then we are already done. 
    % Otherwise, $T'$ intersects the boundary at one edge and its opposing vertex.
    % \todo{MV:But what if this is the neighbor of $T'$ and adding $T'$ back makes the triangle bad?}

@@

We discuss the Sobolev spaces of vector calculus and their transformation behavior. 
Let $\Omega \subseteq \bbR^{3}$ be a bounded open set. 
First, recall that $L^{p}(\Omega)$ is the space of scalar-valued $p$-integrable functions defined on $\Omega$
and that $\bfL^{p}(\Omega) := L^{p}(\Omega)^{n}$ for vector-valued functions with each component in $L^{p}(\Omega)$. 
In the three-dimensional setting, we are particularly interested in the Sobolev vector analysis. 
$W^p(\grad,\Omega) := W^{1,p}(\Omega)$ is the space of scalar-valued $L^{p}(\Omega)$ functions with weak gradients in $\bfL^{p}(\Omega)$, that is,  
\[
    W^p(\grad,\Omega) = \{ u \in L^{p}(\Omega) \suchthat \grad v \in \bfL^{p}(\Omega) \}
\]
The space $\bfW^{p}(\curl,\Omega)$ of vector-valued $\bfL^{p}(\Omega)$ functions with weak curls in $\bfL^{p}(\Omega)$
and the space of vector-valued $\bfL^{p}(\Omega)$ functions with weak divergences in $L^{p}(\Omega)$ are defined as 
\[
    \bfW^{p}(\curl,\Omega) = \{ \bfu \in \bfL^{p}(\Omega) \suchthat \curl \bfu \in \bfL^{p}(\Omega) \},
    \quad 
    \bfW^{p}(\divergence,\Omega) = \{ \bfu \in \bfL^{p}(\Omega) \suchthat \divergence \bfu \in L^{p}(\Omega) \}.
\]
We are interested in transformations of these Sobolev tensor fields from one domain onto another. 
Suppose that $\Omega, \Omega' \subset \mathbb{R}^3$ are open sets and suppose that $\phi: \Omega \to \Omega'$ is a bi-Lipschitz mapping.
The gradient-, curl-, divergence-conforming Piola transformations are the mappings 
$\phi^{\grad}: L^{p}(\Omega') \to L^{p}(\Omega)$,
$\phi^{\curl}: \bfL^{p}(\Omega') \to \bfL^{p}(\Omega)$, 
$\phi^{\divergence}: \bfL^{p}(\Omega') \to \bfL^{p}(\Omega)$,
and
$\phi^{\rm b}: L^{p}(\Omega') \to L^{p}(\Omega)$. 
We define them 
for any $v \in L^{p}(\Omega')$ and $\bfw \in \bfL^{p}(\Omega')$ by setting 
\begin{subequations}\label{eq_definition_piola}
\begin{align}
    \phi^{\grad}(v) &= v \circ \phi, \\
    \phi^{\curl}(\bfw) &= (\Jacobian\phi^{T} \bfw) \circ \phi, \\
    \phi^{\divergence}(\bfw) &= \left( \adj(\Jacobian\phi) \bfw \right)  \circ \phi, \\
    \phi^{\rm b}(v) &= \left( \det(\Jacobian\phi) v \right)  \circ \phi,
\end{align}
\end{subequations}
\todo{MWL: double check the correct sign convention, what is inverse and what is not}
and where $\Jacobian\phi$ is the Jacobian matrix of $\phi$. 
These mappings are invertible. 
%
% TODO: bounds on the Jacobian transform 
%
We use the commutativity relations 
\begin{subequations}\label{eq_piola_commute}
\begin{align}
    \grad (\phi^{\grad} (v)) &= \phi^{\curl}(\grad v), 
    \\
    \curl (\phi^{\curl}(\bfv)) &= \phi^{\divergence}(\curl \bfv), 
    % TODO: watch the sign 
    \\
    \divergence (\phi^{\divergence}(\bfw)) &= \phi^{\rm b}(\divergence \bfw),
\end{align}
\end{subequations}
where $v \in W^{p}(\grad,\Omega)$, $\bfv \in \bfW^{p}(\curl,\Omega)$, and $\bfw \in \bfW^{p}(\divergence,\Omega)$.  
We summarize this as a commuting diagram:
\begin{align*}
    \begin{CD}
        W^{p}(\grad,\Omega') @>{\grad}>> \bfW^{p}(\curl,\Omega') @>{\curl}>> \bfW^{p}(\divergence,\Omega') @>{\divergence}>> L^{p}(\Omega')
        \\
        @VV{\phi^{\grad}}V 
        @VV{\phi^{\curl}}V 
        @VV{\phi^{\divergence}}V 
        @VV{\phi^{\rm b}}V 
        \\
        W^{p}(\grad,\Omega ) @>{\grad}>> \bfW^{p}(\curl,\Omega ) @>{\curl}>> \bfW^{p}(\divergence,\Omega ) @>{\divergence}>> L^{p}(\Omega ).
    \end{CD}
\end{align*}

@@

We write $C^{\infty}\Alt^{k}(\Omega)$ for the space of smooth differential $k$-forms over $\Omega \subseteq \bbR^{n}$,
which is the vector space of smooth mappings from $\Omega$ into $\Alt^{k}(\bbR^{n})$.
The exterior derivative \( \cartan \) is an operator that takes a \( k \)-form \( \omega \in C^{\infty}\Alt^{k}(\Omega) \) 
to a \((k+1)\)-form \( \cartan\omega \in C^{\infty}\Alt^{k+1}(\Omega) \). 
Every \( k \)-form \( \omega \in C^{\infty}\Alt^{k}(\Omega) \) can be written 
\[
    \omega = 
    \sum_{ \sigma \in \Sigma(k,n) } 
    \omega_{\sigma} \, 
    \cartanx^{\sigma(1)} \wedge \cartanx^{\sigma(2)} \wedge \cdots \wedge \cartanx^{\sigma(k)},
\]
where \( \omega_{\sigma} : \Omega \rightarrow \bbR \) are smooth functions.
The exterior derivative \( \cartan\omega \) is defined by
\[
    \cartan\omega = 
    \sum_{ \sigma \in \Sigma(k,n) } 
    \sum_{j=1}^{n} 
    \frac{\partial \omega_{\sigma}}{\partial x^j} 
    \, \cartanx^j \wedge 
    \cartanx^{\sigma(1)} \wedge \cartanx^{\sigma(2)} \wedge \cdots \wedge \cartanx^{\sigma(k)}.
\]
The exterior derivative is linear and nilpotent, which means 
\( \cartan(\cartan\omega) = 0 \) for any \( \omega \in C^{\infty}\Alt^{k}(\Omega) \).
Moreover, it satisfies the Leibniz rule:
\[ 
    \cartan(\omega \wedge \eta) 
    = 
    \cartan\omega \wedge \eta + (-1)^{k} \omega \wedge \cartan\eta, 
    \qquad \omega \in C^{\infty}\Alt^{k}(\Omega), 
    \qquad \eta \in C^{\infty}\Alt^{l}(\Omega).
\]
Let us now turn our attention to Sobolev spaces of differential forms. 
Since the exterior product space $\Alt^{k}(\bbR^{n})$ carries a canonical norm, we can consider the pointwise norms of differential $k$-forms. 
% First, we notice that the % TODO: introduce a norm 
We let $L^{p}\Alt^{k}(\Omega)$ be the space of differential $k$-forms over $\Omega$ with locally integrable coefficients 
such that its pointwise norm is $p$-integrable. 
The exterior derivative is defined in the sense of distributions and we introduce 
\[
    W^p\Alt^{k}(\Omega) 
    := 
    \{ u \in L^{p}\Alt^{k}(\Omega) \suchthat \cartan v \in L^{p}\Alt^{k+1}(\Omega) \}
\]
We observe that $u \in L^p\Alt^{k}(\Omega)$ has weak exterior derivative $f \in L^p\Alt^{k+1}(\Omega)$
if and only if for all $v \in C^{\infty}_{c}\Lambda^{n-k-1}(\Omega)$ we have the integration by parts formula
\begin{align*}
    \int_{\Omega} \cartan v \wedge u
    &=
    (-1)^{k(n-k)+1}
    \int_{\Omega} v \wedge f 
    .
\end{align*}
We let $W^p_{0}\Alt^{k}(\Omega)$ be the closed subspace of those $u \in W^p\Alt^{k}(\Omega)$ 
for which the extension by zero $\tilde u : \bbR^{n} \rightarrow \Alt^{k}(\bbR^{n})$ is a member of $\tilde u \in W^p\Alt^{k}(\bbR^{n})$. 
The space $W^p_{0}\Alt^{k}(\Omega)$ formalizes the idea of differential forms with vanishing boundary trace along the boundary $\partial\Omega$.

@@

\todo{cite Guzman and Salgado at the right spot.}\\
